---
title: "Lego trend analysis"
author: "Kamil Niżnik"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    fig_width: 8
    fig_height: 6
  pdf_document:
    toc: yes
---

<div class="alert alert-info" role="alert">
The data was downloaded on **2023-11-17** from [Rebrickable](https://rebrickable.com). Therefore, some trends may have changed since the report was generated.
</div>

<h2>Executive Summary</h2>
The report analyzes the Lego dataset. The [introduction](#introduction) presents basic statistics about the dataset. The [overall analysis](#overall-analysis) section presents time independent analysis. The [correlation](#correlation) section presents correlation between some variables. The [trend analysis](#trend-analysis) presents time dependent analysis.

**Conclusions and Predictions**

- **Lego part colors**: The most popular Lego colors are black, white and gray (and their shades) [fig. 2.1](#the-most-common-part-colors). This trend does not seem to change [fig. 4.3](#the-most-common-part-colors-over-the-years).
- **Lego minifigures**: Lego minifigures are often designed by inspiration of games (e.g. Minecraft), movies (e.g. Star Wars) or occupation (e.g. policeman or spaceman) [fig. 2.2](#the-most-common-minifigs). It's hard to predict which figures will be popular in the future [fig. 4.4](#the-most-common-minifigs-over-the-years) but the continued design of Lego Collectible Minifigures over the years is expected [fig. 4.2](#the-most-common-themes-over-the-years).
- **Number of Lego sets designed per year**: [Fig. 3.2](#correlation-between-year-and-number-of-sets) shows that there is a high correlation (0.88) between year and number of sets designed. Therefore, probably much more different sets will be designed in the future.
- **Lego set size**: The mean size of Lego set increases over the years whereas the median size stays the same in the recent years [fig. 4.1](#the-set-size-over-the-years). This is because of outliers - there is a high correlation (0.79) between year and size of the biggest set [fig. 3.1](#correlation-between-year-and-max-set-size). In the following years, we can expect single sets larger than World Map, Eiffel Tower or The Ultimate Battle for Chima (the biggest ones).
- **Lego themes**: In the recent years in the Rebrickable database there are plenty of Lego products that are not bricks - especially books and gear (like stickers, umbrellas and other equipment with Lego logo). This trend is more likely to continue [fig 4.2](#the-most-common-themes-over-the-years). Besides, in recent years, Lego Friends, Lego Collectible Minifigures and Lego City sets are often designed. 
- **Lego material**: About 99% of Lego parts are made of plastic [fig 2.5](#the-most-common-part-material). In second place was the rubber. This is a significant environmental problem to solve by Lego.

# Introduction

## Datasets preparation

The following packages are required for robust and reliable report generating.
```{r libraries, message=FALSE, warning=FALSE, error=FALSE}
library(knitr)
library(data.table)
# library(R.utils) - must be installed for 'data.table' gz compression
library(tidyverse)
library(scales)
library(stringr)
library(plotly)
```

```{r opts, include=FALSE}
theme_update(plot.title = element_text(hjust = 0.5))
NUM_PER_CHART <- 12L
IMG_TEMPLATE <- '<img src="%s" alt="No image" height=100 width=100 />'

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}
```

Load the necessary datasets.
```{r read_data, cache=TRUE}
colors_df <- fread("data/colors.csv.gz", colClasses=list(factor=c("is_trans")))
elements_df <- fread("data/elements.csv.gz")
inventories_df <- fread("data/inventories.csv.gz")
inventory_minifigs_df <- fread("data/inventory_minifigs.csv.gz")
inventory_parts_df <- fread("data/inventory_parts.csv.gz", colClasses=list(factor=c("is_spare")))
inventory_sets_df <- fread("data/inventory_sets.csv.gz")
minifigs_df <- fread("data/minifigs.csv.gz")
part_categories_df <- fread("data/part_categories.csv.gz")
part_relationships_df <- fread("data/part_relationships.csv.gz", colClasses=list(factor=c("rel_type")))
parts_df <- fread("data/parts.csv.gz", colClasses=list(factor=c("part_material")))
sets_df <- fread("data/sets.csv.gz")
themes_df <- fread("data/themes.csv.gz")
```

## Datasets summary

**Colors (`r nrow(colors_df)` x `r ncol(colors_df)`)**
```{r data_summary_colors, echo=FALSE}
kable(summary(colors_df), align = "l")
kable(head(colors_df), align = "l")
```

**Elements (`r nrow(elements_df)` x `r ncol(elements_df)`)**
```{r data_summary_elements, echo=FALSE}
kable(summary(elements_df), align = "l")
kable(head(elements_df), align = "l")
```

**Inventories (`r nrow(inventories_df)` x `r ncol(inventories_df)`)**
```{r data_summary_inventories, echo=FALSE}
kable(summary(inventories_df), align = "l")
kable(head(inventories_df), align = "l")
```

**Inventory minifigs (`r nrow(inventory_minifigs_df)` x `r ncol(inventory_minifigs_df)`)**
```{r data_summary_inventory_minifigs, echo=FALSE}
kable(summary(inventory_minifigs_df), align = "l")
kable(head(inventory_minifigs_df), align = "l")
```

**Inventory parts (`r nrow(inventory_parts_df)` x `r ncol(inventory_parts_df)`)**
```{r data_summary_inventory_parts, echo=FALSE}
kable(summary(inventory_parts_df), align = "l")
kable(head(inventory_parts_df), align = "l")
```

**Inventory sets (`r nrow(inventory_sets_df)` x `r ncol(inventory_sets_df)`)**
```{r data_summary_inventory_sets, echo=FALSE}
kable(summary(inventory_sets_df), align = "l")
kable(head(inventory_sets_df), align = "l")
```

**Minifigs (`r nrow(minifigs_df)` x `r ncol(minifigs_df)`)**
```{r data_summary_minifigs, echo=FALSE}
kable(summary(minifigs_df), align = "l")
kable(head(minifigs_df), align = "l")
```

**Part categories (`r nrow(part_categories_df)` x `r ncol(part_categories_df)`)**
```{r data_summary_part_categories, echo=FALSE}
kable(summary(part_categories_df), align = "l")
kable(head(part_categories_df), align = "l")
```

**Part relationships (`r nrow(part_relationships_df)` x `r ncol(part_relationships_df)`)**
```{r data_summary_part_relationships, echo=FALSE}
kable(summary(part_relationships_df), align = "l")
kable(head(part_relationships_df), align = "l")
```

**Parts (`r nrow(parts_df)` x `r ncol(parts_df)`)**
```{r data_summary_parts, echo=FALSE}
kable(summary(parts_df), align = "l")
kable(head(parts_df), align = "l")
```

**Sets (`r nrow(sets_df)` x `r ncol(sets_df)`)**
```{r data_summary_sets, echo=FALSE}
kable(summary(sets_df), align = "l")
kable(head(sets_df), align = "l")
```

**Themes (`r nrow(themes_df)` x `r ncol(themes_df)`)**
```{r data_summary_themes, echo=FALSE}
kable(summary(themes_df), align = "l")
kable(head(themes_df), align = "l")
```

# Overall Analysis

## The most common part colors
```{r analysis_part_colors, echo=FALSE, message=FALSE}
inventory_parts_df %>%
  inner_join(colors_df, by=c("color_id" = "id")) %>%
  mutate(rgb = paste("#", rgb, sep="")) %>%
  select(rgb, name) %>%
  group_by(rgb, name) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART) %>%
  ggplot(aes(x = reorder(name, count), y = count, fill = rgb, color = "black")) +
    geom_bar(stat="identity") +
    scale_fill_identity() +
    scale_color_identity() +
    xlab("Color") +
    ylab("The number of parts") +
    scale_y_continuous(labels = label_comma()) +
    ggtitle("The most common parts colors") +
    coord_flip()
```

## The most common minifigs
```{r analysis_minifigs, echo=FALSE, message=FALSE}
inventory_minifigs_df %>%
  inner_join(inventories_df, by=c("inventory_id" = "id")) %>%
  select(fig_num) %>%
  inner_join(minifigs_df) %>%
  group_by(fig_num, name) %>%
  summarize(count = n()) %>%
  mutate(name=str_trunc(name, 25)) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART) %>%
  ggplot(aes(x = reorder(name, count), y = count)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Minifig name") +
    ylab("The number of sets") +
    coord_flip() +
    ggtitle("The most common minifigs")
```

```{r analysis_minifigs_table, echo=FALSE, message=FALSE}
inv_figs_table <- inventory_minifigs_df %>%
  inner_join(inventories_df, by=c("inventory_id" = "id")) %>%
  select(fig_num) %>%
  inner_join(minifigs_df) %>%
  group_by(name, img_url) %>%
  summarize(count = n()) %>%
  mutate(img_url = sprintf(IMG_TEMPLATE, img_url)) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART)

kable(list(head(inv_figs_table, n=NUM_PER_CHART %/% 2L),
           tail(inv_figs_table, n=NUM_PER_CHART %/% 2L)),
      col.names = c("name", "image", "number"))
rm(inv_figs_table)
```

## The most common parts
```{r analysis_parts, echo=FALSE, message=FALSE}
parts_df %>%
  rename(part_name = name) %>%
  inner_join(inventory_parts_df) %>%
  inner_join(inventories_df, by = c("inventory_id" = "id")) %>%
  inner_join(sets_df, by = c("set_num" = "set_num")) %>%
  group_by(part_num, part_name) %>%
  summarise(count = n_distinct(set_num)) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART) %>%
  mutate(part_name=str_trunc(part_name, 25)) %>%
  ggplot(aes(x = reorder(part_name, count), y = count)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Part name") +
    ylab("The number of sets") +
    scale_y_continuous(labels = label_comma()) +
    coord_flip() +
    ggtitle("The most common parts")
```

```{r analysis_parts_table, echo=FALSE, message=FALSE}
parts_table <- parts_df %>%
  rename(part_name = name) %>%
  inner_join(inventory_parts_df) %>%
  inner_join(inventories_df, by = c("inventory_id" = "id")) %>%
  inner_join(sets_df, by = c("set_num" = "set_num")) %>%
  group_by(part_num, part_name) %>%
  summarise(count = n_distinct(set_num)) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART) %>%
  ungroup() %>%
  inner_join(inventory_parts_df %>% distinct(part_num, .keep_all = TRUE)) %>%
  mutate(img_url = sprintf(IMG_TEMPLATE, img_url)) %>%
  select(part_name, img_url, count)

kable(list(head(parts_table, n=NUM_PER_CHART %/% 2L),
           tail(parts_table, n=NUM_PER_CHART %/% 2L)),
      col.names = c("name", "image", "number"))
rm(parts_table)
```

## The most common part categories
```{r analysis_part_categories, echo=FALSE, message=FALSE}
parts_df %>%
  rename(part_name = name) %>%
  inner_join(part_categories_df, by = c("part_cat_id" = "id"))%>%
  rename(cat_name = name) %>%
  inner_join(inventory_parts_df) %>%
  inner_join(inventories_df, by = c("inventory_id" = "id")) %>%
  inner_join(sets_df, by = c("set_num" = "set_num")) %>%
  group_by(cat_name) %>%
  summarise(count = n_distinct(set_num)) %>%
  arrange(desc(count)) %>%
  head(n=NUM_PER_CHART) %>%
  mutate(cat_name=str_trunc(cat_name, 25)) %>%
  ggplot(aes(x = reorder(cat_name, count), y = count)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Part category name") +
    ylab("The number of sets") +
    scale_y_continuous(labels = label_comma()) +
    coord_flip() +
    ggtitle("The most common part categories")
```

## The most common part material
```{r analysis_part_material, echo=FALSE, message=FALSE}
parts_df %>%
  inner_join(inventory_parts_df) %>%
  group_by(part_material) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(count = (count / sum(count)) * 100) %>%
  head(n=NUM_PER_CHART) %>%
  ggplot(aes(x = reorder(part_material, count), y = count)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Material") +
    ylab("[%] of parts") +
    scale_y_continuous(labels = label_comma()) +
    coord_flip() +
    ggtitle("The most common part material")
```

## The most complex sets
```{r analysis_complex_sets, echo=FALSE, message=FALSE}
sets_df %>%
  arrange(desc(num_parts)) %>%
  head(n=NUM_PER_CHART) %>%
  ggplot(aes(x = reorder(name, num_parts), y = num_parts)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Set name") +
    ylab("The number of parts") +
    scale_y_continuous(labels = label_comma()) +
    coord_flip() +
    ggtitle("The most complex sets")
```

```{r analysis_complex_sets_table, echo=FALSE, message=FALSE}
comp_sets_table <- sets_df %>%
  arrange(desc(num_parts)) %>%
  head(n=NUM_PER_CHART) %>%
  mutate(img_url = sprintf(IMG_TEMPLATE, img_url)) %>%
  select(name, img_url, num_parts)

kable(list(head(comp_sets_table, n=NUM_PER_CHART %/% 2L),
           tail(comp_sets_table, n=NUM_PER_CHART %/% 2L)),
      col.names = c("name", "image", "number"))
rm(comp_sets_table)
```

## The most common themes
```{r analysis_themes, echo=FALSE, message=FALSE}
themes_df %>%
  mutate(parent_id = ifelse(is.na(parent_id), id, parent_id))%>%
  rename(theme = name) %>%
  inner_join(sets_df, by = c("id" = "theme_id")) %>%
  group_by(parent_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  inner_join(themes_df, by = c("parent_id" = "id")) %>%
  distinct(name, .keep_all = TRUE) %>%
  select(name, count) %>%
  head(n=NUM_PER_CHART) %>%
  ggplot(aes(x = reorder(name, count), y = count)) +
    geom_bar(stat="identity", fill="steelblue") +
    xlab("Theme") +
    ylab("The number of sets") +
    scale_y_continuous(labels = label_comma()) +
    coord_flip() +
    ggtitle("The most common themes")
```

# Correlation

## Correlation between year and max set size
```{r corr_year_size, echo=FALSE, message=FALSE}
(sets_df %>%
  filter(year >= 1950, year <= 2022) %>%
  select(year, num_parts) %>%
  group_by(year) %>%
  summarise(max = max(num_parts)) %>%
  ggplot(aes(x = year, y = max)) +
    geom_point() +
    geom_smooth() +
    scale_y_continuous(labels = label_comma()) +
    xlab("Year") +
    ylab("The maximum set size (number of parts)") +
    ggtitle("Correlation between year and max set size")) %>%
  ggplotly
```


**Correlation matrix**
```{r corr_year_size_mat, echo=FALSE, message=FALSE}
sets_df %>%
  filter(year >= 1950, year <= 2022) %>%
  select(year, num_parts) %>%
  group_by(year) %>%
  summarise(max = max(num_parts)) %>%
  cor %>%
  kable
```

## Correlation between year and number of sets
```{r corr_year_set_num, echo=FALSE, message=FALSE}
(sets_df %>%
  filter(year >= 1950, year <= 2022) %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = year, y = count)) +
    geom_point() +
    geom_smooth() +
    scale_y_continuous(labels = label_comma()) +
    xlab("Year") +
    ylab("The number of sets") +
    ggtitle("Correlation between year and number of sets")) %>%
  ggplotly
```


**Correlation matrix**
```{r corr_year_set_num_mat, echo=FALSE, message=FALSE}
sets_df %>%
  filter(year >= 1950, year <= 2022) %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  cor %>%
  kable
```

# Trend Analysis

## The set size over the years
```{r trend_size_sets, echo=FALSE, message=FALSE}
(sets_df %>%
  filter(year >= 1950, year <= 2022) %>%
  group_by(year) %>%
  summarise(median = median(num_parts), mean = mean(num_parts), max = max(num_parts)) %>%
  select(year, median, mean, max) %>%
  gather(key = "type", value = "value", -year) %>%
  mutate(value = round(value, 2)) %>%
  ggplot(aes(x = year, y = value, color = type)) +
    geom_line() +
    geom_point() +
    xlab("Year") +
    ylab("The number of parts") +
    scale_y_continuous(labels = label_comma()) +
    ggtitle("The set size over the years") +
    facet_grid(type~., scales="free_y") +
    theme(legend.position = "none")) %>%
  ggplotly
```

## The most common themes over the years
```{r trend_themes, echo=FALSE, message=FALSE}
themes_df %>%
    mutate(parent_id = ifelse(is.na(parent_id), id, parent_id))%>%
    rename(theme = name) %>%
    inner_join(sets_df, by = c("id" = "theme_id")) %>%
    group_by(year, parent_id) %>%
    summarise(count = n()) %>%
    arrange(desc(count)) %>%
    inner_join(themes_df, by = c("parent_id" = "id")) %>%
    distinct(name, .keep_all = TRUE) %>%
    select(year, name, count) %>%
    group_by(year) %>%
    slice_max(order_by = count, n = 5, with_ties = FALSE) %>%
    filter(year %in% c(1980, 1990, 2000, 2010, 2020)) %>%
    ggplot(aes(x = reorder_within(name, count, year), y = count, fill = factor(year))) +
      geom_bar(stat="identity") +
      xlab("Theme") +
      ylab("The number of sets") +
      coord_flip() +
      ggtitle("The most common themes") +
      facet_grid(year~., scales = "free_y") +
      theme(legend.position = "none") +
      scale_fill_brewer(palette = "Set2") +
      scale_x_reordered()
```

```{r trend_themes_sec, echo=FALSE, message=FALSE, fig.height=10}
(themes_df %>%
    mutate(parent_id = ifelse(is.na(parent_id), id, parent_id))%>%
    rename(theme = name) %>%
    inner_join(sets_df, by = c("id" = "theme_id")) %>%
    group_by(year, parent_id) %>%
    summarise(count = n()) %>%
    arrange(desc(count)) %>%
    inner_join(themes_df, by = c("parent_id" = "id")) %>%
    distinct(name, .keep_all = TRUE) %>%
    select(year, name, count) %>%
    filter(name %in% c("Gear", "Duplo", "Books", "Star Wars", "Collectible Minifigures"), year >= 1980, year <= 2022) %>%
    ggplot(aes(x = year, y = count, color = name)) +
      geom_line() +
      geom_point() +
      xlab("Year") +
      ylab("The number of sets") +
      ggtitle("The most common themes") +
      facet_grid(name~., scales = "free_y") +
      scale_color_brewer(palette = "Set2") +
      theme(legend.position = "none")) %>% ggplotly
```


## The most common part colors over the years
```{r trend_colors, echo=FALSE, message=FALSE}
inventory_parts_df %>%
  inner_join(colors_df, by=c("color_id" = "id")) %>%
  mutate(rgb = paste("#", rgb, sep="")) %>%
  rename(color_name = name) %>%
  select(c(inventory_id, part_num, color_name, rgb)) %>%
  inner_join(inventories_df, by = c("inventory_id" = "id")) %>%
  inner_join(sets_df) %>%
  select(year, color_name, rgb) %>%
  group_by(year, color_name, rgb) %>%
  summarise(count = n()) %>%
  group_by(year) %>%
  slice_max(order_by = count, n = 5, with_ties = FALSE) %>%
  filter(year %in% c(1980, 1990, 2000, 2010, 2020)) %>%
  ggplot(aes(x = reorder_within(color_name, count, year), y = count, fill = rgb, color="black")) +
      geom_bar(stat="identity") +
      xlab("Color") +
      ylab("The number of inventory parts") +
      coord_flip() +
      ggtitle("The most common colors") +
      facet_grid(year~., scales = "free_y") +
      theme(legend.position = "none") +
      scale_x_reordered() +
      scale_fill_identity() +
      scale_color_identity() +
      scale_y_continuous(labels = label_comma())
```

## The most common minifigs over the years
```{r trend_minifigs, echo=FALSE, message=FALSE}
inventory_minifigs_df %>%
  inner_join(inventories_df, by=c("inventory_id" = "id")) %>%
  inner_join(minifigs_df) %>%
  select(fig_num, set_num, name) %>%
  inner_join(sets_df, by = c("set_num" = "set_num")) %>%
  select(year, name.x) %>%
  rename(name = name.x) %>% 
  group_by(year, name) %>%
  summarise(count = n()) %>%
  slice_max(order_by = count, n = 4, with_ties = FALSE) %>%
  filter(year %in% c(2020, 2021, 2022, 2023)) %>%
  mutate(name=str_trunc(name, 25)) %>%
  ggplot(aes(x = reorder_within(name, count, year), y = count, fill = factor(year))) +
      geom_bar(stat="identity") +
      xlab("Minifig name") +
      ylab("The number of sets") +
      coord_flip() +
      ggtitle("The most common minifigs") +
      facet_grid(year~., scales = "free_y") +
      theme(legend.position = "none") +
      scale_fill_brewer(palette = "Set2") +
      scale_x_reordered() +
      scale_y_continuous(labels = label_comma())
```